### 2015-04-16 implementation-environment.tex
GLL edits

%{\bf Reproducibility.}
%\paragraph{Reproducibility}
Spack is intended to build a consistent HPC stack for our multi-user
environment, and reproducible builds are one of our design goals.
Experience at LLNL has shown that it is vexingly difficult to reproduce
a build manually.
%
Many packages used at LLNL have a profusion of build options, and specifying them 
correctly often requires tedious experimentation.  This is due to lack of
build standards and to the diversity of HPC environments.  
For example, in some packages that depend on the {\tt Silo} library,
the {\tt ---with-silo} parameter takes a path to f{\tt Silo}'s prefix.
In others, it takes the {\tt include} and {\tt lib} subdirectories,
separated by a comma.
The {\tt install()} method in Spack's package files allows us to record
precise build incantations for later reuse.

%{\bf Environment isolation.}
\paragraph{Environment isolation}
In addition to command line issues, we
frequently encounter errors due to inconsistencies between the environment of
the package installer and the package user.
%
For example, there are two versions of the {\tt libelf} library used by
LLNL performance tools. One is distributed with RedHat Linux, but another
publicly available version has the same API but an incompatible ABI.
Failure to specify the right version at build time has caused many
inexplicable crashes.
%
Spack manages the build environment by running {\tt install} invocation
in a new process.  It helps packages find dependencies 
correctly, by setting
{\tt PATH}, {\tt PKG\_CONFIG\_PATH}, {\tt CMAKE\_PREFIX\_PATH}, and
{\tt LD\_LIBRARY\_PATH} to include the dependencies of the current build.
These variables are commonly used by build systems to locate dependencies,
and setting them helps to ensure that incorrect libraries are not detected.
The isolated build environment also gives package authors 
free reign to set build-specific environment variables without interfering
with other packages.

@@

Spack's compiler wrappers have a number of useful effects.  First, they allow
Spack to transparently parameterize the compiler for most builds. 
Spack users to easily request a build with a new compiler swapped in.
Second, they enforce the use of {\tt RPATHs} in
installed binaries.  This causes applications built by Spack to run correctly
{\it regardless of the environment}.  Third, because compiler wrappers add 
header and library search paths for dependencies, header and library detection
tests run by most build systems succeed automatically, {\it without}
the need to use special arguments for nonstandard locations.  {\tt configure}
commands in Spack's {\tt install} function can have fewer arguments, and can
be written as they would be for system installs.  This reduces complexity
for package maintainers and enforces consistent, reproducible
build policies across packages.  Finally, because Spack has control over the 
wrappers, package authors can programmatically filter the compiler flags
used by software build systems, a useful last resort when porting to
bleeding-edge platforms or new, esoteric compilers.

>>>>

%{\bf Reproducibility.}
%\paragraph{Reproducibility}
Spack is intended to build a consistent HPC stack for our multi-user
environment, and reproducible builds are one of our design goals.
Experience at LLNL has shown that it is vexingly difficult to reproduce
a build manually.
%
Many packages used at LLNL have a profusion of build options, and specifying them 
correctly often requires tedious experimentation.  This is due to lack of
build standards and to the diversity of HPC environments.  
For example, in some packages that depend on the {\tt Silo} library,
the {\tt ---with-silo} parameter takes a path to {\tt Silo}'s installation prefix.
In others, it takes the {\tt include} and {\tt lib} subdirectories,
separated by a comma.
The {\tt install()} method in Spack's package files allows us to record
precise build incantations for later reuse.

%{\bf Environment isolation.}
\paragraph{Environment isolation}
In addition to command-line issues, we
frequently encounter errors due to inconsistencies between the environment of
the package installer and the package user.
%
For example, there are two versions of the {\tt libelf} library used by
LLNL performance tools. One is distributed with RedHat Linux, while another
publicly available version has the same API but an incompatible ABI.
Failure to specify the right version at build time has caused many
inexplicable crashes.
%
Spack manages the build environment by running the {\tt install} invocation
in a new process.  It helps packages find dependencies 
correctly, by setting
{\tt PATH}, {\tt PKG\_CONFIG\_PATH}, {\tt CMAKE\_PREFIX\_PATH}, and
{\tt LD\_LIBRARY\_PATH} to include the dependencies of the current build.
These variables are commonly used by build systems to locate dependencies,
and setting them helps to ensure that incorrect libraries are detected.
The isolated build environment also gives package authors 
free reign to set build-specific environment variables without interfering
with other packages.

@@

Spack's compiler wrappers have a number of useful effects.  First, they allow
Spack to transparently parameterize the compiler for most builds and for
Spack users to easily request a build with a new compiler swapped in.
Second, they enforce the use of {\tt RPATHs} in
installed binaries.  This causes applications built by Spack to run correctly
{\it regardless of the environment}.  Third, because compiler wrappers add 
header and library search paths for dependencies, header and library detection
tests run by most build systems succeed automatically, {\it without}
the need to use special arguments for nonstandard locations.  {\tt configure}
commands in Spack's {\tt install} function can have fewer arguments, and can
be written as they would be for system installs.  This reduces complexity
for package maintainers and enforces consistent, reproducible
build policies across packages.  Finally, because Spack has control over the 
wrappers, package authors can programmatically filter the compiler flags
used by software build systems, a useful last resort when porting to
bleeding-edge platforms or new, esoteric compilers.

### 2015-04-16 implementation-packages.tex
GLL edits

Figure~\ref{fig:mpileaks} shows the package for \mpileaks, an LLNL-developed
tool developed for finding leaks in MPI programs.
Inside the {\tt MpiLeaks} class, the package provides a text description
and a homepage, as well as 
a download URL.  Next, two {\tt version} directives identify known versions
of the package, and MD5 checksums ensure they can be downloaded safely.
Below this, three {\tt depends\_on}
directives indicate prerequisites that must be installed before \mpileaks.
Last, each package defines an {\tt install()} method, which contains the
commands used to build the package.  Spack's DSL allows shell
commands to be invoked as Python functions. Here, the {\tt install()} 
method invokes the familiar {\tt configure}, {\tt make}, and
{\tt make install} build commands, as a shell script would.

>>>>

Figure~\ref{fig:mpileaks} shows the package for \mpileaks, an LLNL-developed
tool for finding leaks in MPI programs.
Inside the {\tt MpiLeaks} class, the package provides a text description
and a homepage, as well as 
a download URL.  Next, two {\tt version} directives identify known versions
of the package, and MD5 checksums ensure they can be downloaded safely.
Below this, two {\tt depends\_on}
directives indicate prerequisites that must be installed before \mpileaks.
Last, each package defines an {\tt install()} method, which contains the
commands used to build the package.  Spack's DSL allows shell
commands to be invoked as Python functions. Here, the {\tt install()} 
method invokes the familiar {\tt configure}, {\tt make}, and
{\tt make install} build commands, as a shell script would.

### 2015-04-16 implementation-specs.tex
GLL edits

{\bf Versions.}
Versions are specified by supplying {\tt @<version>} after the package name (line 2). 
Versions can be precise (e.g. {\tt @4.5.1}) or 
they can be ambiguous (e.g., @4) or a colon-separated, inclusive range (e.g., @4.4:4.5.1),
and Spack will pick an available compiler from the specified range.
The package in Figure~\ref{fig:mpileaks} lists two ``safe'' versions with checksums, but
in our experience users frequently want bleeding-edge versions.  Package managers
frequently lag behind the latest releases. 
Spack has a capability to extrapolate URLs from versions,
using the package's {\tt url} attribute as a model\footnote{This works
for packages with consistently named URLs}.  The user can request a specific
version on the command line, even if it is unknown to Spack,
and Spack will attempt to install it.  Spack also uses the same
model to scrape webpages and find new versions as they become available.

@@

{\bf Variants.}
To handle build-time options like debug flags and optional components, spack
allows named flags, or {\it variants} to be associated with packages.  
In the example (line 4), {\tt mpileaks} can be built with a {\tt debug} option by
adding {\tt +debug}.  The option can be similarly disabled with {\tt -debug}.  
This allows package authors to provide package-specific build options without 
having to extend Spack.  We use names for these to simplify the spec
syntax, and to prevent Spack's configuration space from becoming too fine-grained.
It would be violate our goal of conciseness, for example, to include detailed 
compiler flags in the spec syntax, but known sets of flags can be exported using
named variants.

@@

{\bf Dependencies.}
Lines 6 and 7 in the table show the key feature that enables Spack's flexibility.
The user can specify all of the above information not only for the package being
installed, but also for its {\it dependencies}.  To do this, the user needs only supply 
\verb|^| and the dependency's name.  If we need to build a new version with a specific
version of {\tt mvapich2}, we can simply add, e.g., \verb|^mvapich2@1.9|
to the spec, and it will build with that MPI version instead of the default.
This can also be done for multiple libraries in the same spec (line 7).  
As mentioned, Spack guarantees that there is only one version of any package in 
the spec DAG.  So, within the same DAG, each dependency can be uniquely identified by 
only its package name.  This frees the user from having to worry about DAG connectivity;
the user only needs to know that a package depends, somehow, indirectly on {\tt mvapich}
to add this constrain
uniquely identifies a dependency, and dependency constraints can be specified in any
order.

>>>>

{\bf Versions.}
Versions are specified by supplying {\tt @<version>} after the package name (line 2). 
Versions can be precise (e.g. {\tt @4.5.1}) or 
they can be ambiguous (e.g., @4) or a colon-separated, inclusive range (e.g., @4.4:4.5.1),
and Spack will pick an available version from the specified range.
The package in Figure~\ref{fig:mpileaks} lists two ``safe'' versions with checksums, but
in our experience users frequently want bleeding-edge versions.  Package managers
frequently lag behind the latest releases. 
Spack has a capability to extrapolate URLs from versions,
using the package's {\tt url} attribute as a model\footnote{This works
for packages with consistently named URLs}.  The user can request a specific
version on the command line, even if it is unknown to Spack,
and Spack will attempt to install it.  Spack also uses the same
model to scrape webpages and find new versions as they become available.

@@

{\bf Variants.}
To handle build-time options like debug flags and optional components, spack
allows named flags, or {\it variants} to be associated with packages.  
In the example (line 4), {\tt mpileaks} can be built with a {\tt debug} option by
adding {\tt +debug}.  The option can be similarly disabled with {\tt -debug}.  
This allows package authors to provide package-specific build options without 
having to extend Spack.  We use names for these to simplify the spec
syntax, and to prevent Spack's configuration space from becoming too fine-grained.
It would violate our goal of conciseness, for example, to include detailed 
compiler flags in the spec syntax, but known sets of flags can be exported using
named variants.

@@

{\bf Dependencies.}
Lines 6 and 7 in the table show the key feature that enables Spack's flexibility.
The user can specify all of the above information not only for the package being
installed, but also for its {\it dependencies}.  To do this, the user needs only supply 
\verb|^| and the dependency's name.  If we need to build a new version with a specific
version of {\tt mvapich2}, we can simply add, e.g., \verb|^mvapich2@1.9|
to the spec, and it will build with that MPI version instead of the default.
This can also be done for multiple libraries in the same spec (line 7).  
As mentioned, Spack guarantees that there is only one version of any package in 
the spec DAG.  So, within the same DAG, each dependency can be uniquely identified by 
only its package name.  This frees the user from having to worry about DAG connectivity;
the user only needs to know that a package depends, somehow, indirectly on {\tt mvapich}
to add this constraint
uniquely identifies a dependency, and dependency constraints can be specified in any
order.
